---
title: "Sydney Beach Quality and Safety Assessment"
author: "Group: chocolate            Jingyi Wang"
format: 
  revealjs:
    theme: simple
    slide-number: true
    toc: true
    toc-depth: 2
execute:
  echo: false
  warning: false
editor: visual
---
```{r}
library(tidyverse)
library(knitr)
library(viridis)
library(viridisLite)
```


## Executive Summary

-   Analyzed 10 years of Sydney beach data (2015–2025)\
-   Focused on **enterococci levels** (bacteria in water)\
-   Found **some beaches highly polluted**, especially Wentworth Falls Lake\
-   **Rainfall strongly increases** bacteria levels\
-   Urge for **real-time water quality monitoring** and public alerts

------------------------------------------------------------------------

## Introduction

-   Sydney beaches are popular but sometimes **unsafe** for swimming\
-   Enterococci bacteria levels are key indicators of **water pollution**\
-   **Research Questions**:
    -   Which beaches are most polluted?
    -   Does **rainfall** increase bacteria levels?

------------------------------------------------------------------------

## Dataset Description

-   **Water Quality** (Enterococci CFU/100ml)\
-   **Weather Data** (Daily rainfall, temperature)\
-   Time range: **2015–2025**\
-   Data Source: [TidyTuesday, May 2025](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-20/readme.md)

```{r}
water_quality <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-20/water_quality.csv')
weather <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-20/weather.csv')

start_date <- as.Date("2015-04-28")
end_date <- as.Date("2025-04-28")

water_quality <- water_quality %>%
  filter(date >= start_date & date <= end_date)

weather <- weather %>%
  filter(date >= start_date & date <= end_date)
```

------------------------------------------------------------------------

## Methodology (1/2): Worst Beaches

-   Calculated **average enterococci levels** by beach

-   Identified **Top 10 most polluted** sites

-   Used bar chart to show comparison

```{r}
top_beaches <- water_quality %>%
  group_by(swim_site) %>%
  summarise(avg_enterococci = mean(enterococci_cfu_100ml, na.rm = TRUE)) %>%
  arrange(desc(avg_enterococci)) %>%
  slice(1:10)
```

```{r}
#| label: fig-bar_chart
#| fig-cap: "Top 10 Beaches with Worst Water Quality"
#| echo: false

ggplot(top_beaches, aes(x = reorder(swim_site, avg_enterococci), 
                        y = avg_enterococci,
                        fill = reorder(swim_site,
                                       avg_enterococci))) +  
  
  geom_col() +  
  
  scale_fill_viridis(discrete = TRUE, 
                     option = "B", 
                     direction = -1) +  
  
  coord_flip() +
  
  labs(title = "Top 10 Beaches with Worst Water Quality",
       subtitle = "Based on average Enterococci levels",
       x = "Beach",
       y = "Average Enterococci (CFU/100ml)",
       caption = "Data source: TidyTuesday") +
  
  theme_minimal(base_size = 10) +
  
  theme(
    plot.title = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "gray40"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    legend.position = "none",  # Remove legend
    panel.grid.major.y = element_blank(),  # Remove horizontal grid lines
    plot.caption = element_text(color = "gray50", hjust = 0)
  )

```

------------------------------------------------------------------------

## Methodology (2/2): Rainfall & Water Quality

-   Merged water and weather datasets by date

-   Grouped rainfall:

    -   `0 mm`, `0–10 mm`, `>10 mm`

-   Calculated average bacteria level per group

-   Log10 boxplot for better visibility

```{r}
# In weather dataset, both latitude and longitude have only unique values.
weather <- weather %>%
  select(-latitude, -longitude)

# Merge water and weather datasets by date
water_weather <- water_quality %>%
  left_join(weather, by = "date")

```

```{r}
water_weather <- water_weather %>%
  mutate(rain_group = case_when(
    precipitation_mm == 0 ~ "0 mm",
    precipitation_mm <= 10 ~ "0–10 mm",
    TRUE ~ ">10 mm"
  ))

```

------------------------------------------------------------------------

## Bacteria Level by Rainfall

::: {.columns}
::: {.column width="35%"}


```{r}
#| label: tbl-rain_table
#| tbl-cap: "Average Enterococci by Rain Group"

rain_table <- water_weather %>%
  group_by(rain_group) %>%
  summarise(mean_enterococci = mean(enterococci_cfu_100ml, na.rm = TRUE))

kable(rain_table, format = "html", table.attr = "style='font-size: 55%;'")

```

:::
::: {.column width="55%"}

```{r}
#| label: fig-boxplot
#| fig-cap: "Boxplot of Enterococci Levels by Rainfall Group"

ggplot(water_weather, 
       aes(x = rain_group, 
           y = enterococci_cfu_100ml, 
           fill = rain_group)) +  
  geom_boxplot(
    width = 0.6,                
    color = "black",            
    size = 0.5,                 
    outlier.alpha = 0.5   # Outlier transparency
  ) +
  scale_y_log10() +
  scale_fill_manual(
    values = c("0 mm" = "#FDE725",    
               "0–10 mm" = "#7AD151",
               ">10 mm" = "#31688E"), 
    breaks = c("0 mm", "0–10 mm", ">10 mm")) +
  labs(
    title = "Enterococci Levels by Rainfall Group",
    subtitle = "Log-transformed bacteria concentration (CFU/100ml)",
    x = "Rainfall Group",
    y = "Enterococci (CFU/100ml, log10)",
    caption = "Color intensity reflects rainfall amount. Darker colors indicate higher rainfall."
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(color = "gray40"),
    axis.title = element_text(face = "bold"),
    legend.position = "none",     # Remove legend
    panel.grid.major.x = element_blank(), # Remove vertical grid lines
    panel.grid.minor = element_blank())

```

:::
:::

------------------------------------------------------------------------

## Result & Conclusion

**Result:**
```{=html}
<div style="font-size: 70%;">
```
-   The beach with the highest average pollution level is Wentworth Falls Lake - Jetty

-   It suggests that certain beaches consistently have poor water quality from 2015 to 2025

-   Beaches with rainfall greater than 10mm had significantly higher enterococci levels compared to those with 0mm rainfall.

-   These findings are consistent with prior research, including Wade et al.(2012) and  Soller et al. (2010).
```{=html}
</div>
```
**Conclusion:**
```{=html}
<div style="font-size: 70%;">
```
- Some beaches in Sydney → Dangerous → The risk of getting sick

- Heavy rainfall → Beach water becomes unsafe
```{=html}
</div>
```

```{=html}
<div style="font-size: 80%;">
```
**Many beaches are not always safe for swimming, especially after rain. **
```{=html}
</div>
```
--- 

## Recommendations & References

**Recommendations:**

- Increase testing frequency

- Publish data publicly in real time

- Post warning signs after rain

- Use forecast + data to build early warning systems

**Key References:**

- Wade et al. (2012)

- Soller et al. (2010)
